---
title: "Report_CohortAnalysis"
author: "Nils Gandlau"
date: "2 12 2020"
output: html_document
---

## Cohort Analysis of Customer Churn

```{r}
library(data.table)
library(tidyverse)
library(magrittr)
```

```{r}
source("functions.R")
translog <- read_csv("../data/fake_data.csv") %>%
  setDT()
translog <- AddAcquisitionTimestamp(translog, "orderTimestamp", "customerId")
```

### Goal

Explore and understand the current stand of customer churn. Through cohort analysis, we can investigate the age, period, and cohort effects that make up customer churn.

## 1. Visual Exploration

**User Input 0**: Choose the metric he wants to look at.

**User Input 1**: Choose among *monthly, quarterly*, or *yearly* cohorts using `AddMonthlyCohorts(), AddQuarterlyCohorts(), AddYearlyCohorts()`

**User Input 2**: Choose whether to display the metric *1) absolute 2) relative to previous period or 3) relative to acquisition period. (TODO)*

**User Input 3**: Choose whether to display *1) Cohort-Age* or *2) Cohort-Period* View specifying `view = "CA"` or `view = "CP"` in `PlotCohortTable()`.

UserInput 4: Choose the type of plot he wants to display. Either `PlotCohortTable()` or `PlotCohortLinechart()`.

```{r}
CohortMetric <- function(data, metric = "nCustomers"){
  if (metric == "nCustomers"){
    CohortCustomers(data)  
  } else if (metric == "revenue"){
    CohortRevenue(data)
  } else if (metric == "LTV30") {
    CohortLTV(data, days = 30, summarise = "mean")
  } else if (metric == "LTV60") {
    CohortLTV(data, days = 60, summarise = "mean")
  } else if (metric == "LTV30/CAC") {
    CohortLTVCAC(data, days = 30, summarise = "mean")
  } else if (metric == "LTV60/CAC") {
    CohortLTVCAC(data, days = 60, summarise = "mean")
  }
}
```

```{r}
CohortType <- function(data, cohortType = "m",
                       orderTimestampCol = "orderTimestamp",
                       acqTimestampCol = "acqTimestamp",
                       customerIdCol = "customerId"){
  # Args:
  #   cohortType %in% c("m", "q", "y") (corresponding to month, quarter, year)
  if (cohortType == "m"){
    AddMonthlyCohorts(data, orderTimestampCol, acqTimestampCol, customerIdCol)
  } else if (cohortType == "q"){
    AddQuarterlyCohorts(data, orderTimestampCol, acqTimestampCol, customerIdCol)
  } else if (cohortType == "y"){
    AddYearlyCohorts(data, orderTimestampCol, acqTimestampCol, customerIdCol)
  }
}
```

```{r}
# User Inputs
cohortType <- "m"
metric <- "nCustomers"
plotType <- "table"
view <- "CP"

# Output
translog %>% 
  CohortType(cohortType = cohortType) %>% 
  CohortMetric(metric = metric) %>% 
  PlotCohort(plotType = plotType, view = view)
```

```{r}
# User Inputs
cohortType <- "q"
metric <- "revenue"
plotType <- "linechart"
view <- "CA"

# Output
translog %>% 
  CohortType(cohortType = cohortType) %>% 
  CohortMetric(metric = metric) %>% 
  PlotCohort(plotType = plotType, view = view)
```

## 2. APC Regression

...
