---
title: "Customer Churn Analysis"
author: "Nils Gandlau"
date: "25 11 2020"
output: 
  rmdformats::html_clean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = FALSE)
```

## Research Questions

**Question:** Who is churning? Which are common characteristics among customers who stay longer?

**Implications:** Deciding which advertisement channels is the best to acquire long-term customers. Discovery of customer characteristics associated with long-term relationships allows you to better choose your target groups when advertising.

------------------------------------------------------------------------

**Question:** Do coupons improve customer churn?

**Implications:** Decide whether coupons are a investment with a positive return.

------------------------------------------------------------------------

**Question:** How long are customers staying?

**Implications:** Understand the long-term growth capabilities of your business.

## Prerequisites

-   Definition of Churn: Customers who have not bought in the previous 1 month. (chosen by the user)

-   Choosing covariates for the regression (outsourced to us)

-   Choosing the right model (outsourced to us)

## Data

```{r}
library(glue)
library(data.table)
data <- fread(file = glue("../data/fake_data.csv"))
```

We define a *churned customer* as a customer who has not bought anything in the last 2 months.

```{r}
#DefineChurners <- function(data, notBoughtForDays = 60){
data[, lastPurchaseDate := max(orderTimestamp), by = customerId]
data[, endOfStudyDate := max(orderTimestamp)]
data[, isChurn := ifelse(endOfStudyDate - lastPurchaseDate >= 60, 1, 0)]

# Length of relationship
data[, survTime := endOfStudyDate - min(orderTimestamp), by = customerId]
data[isChurn == 1, 
     survTime := max(orderTimestamp) - min(orderTimestamp),
     by = customerId]
#}
```

## Who is churning?

**Method:** Survival Analysis; Cox Model; Parametric Proportional Hazards Model

**User:** Filter for min. number of purchases (e.g. remove one-time buyers who would bias the results)

```{r echo=FALSE, include=TRUE}
library(survival)
cox <- coxph(Surv(survTime, isChurn) ~ acqChannel + age + salary, data = data)
summary(cox)
```

**🗻Insights**

-   Customers with a high income are associated with having a 1.53 times longer relationships than those with low or medium income. \
    =\> You should target those.

-   Young customers stay longer than older ones. In fact, a 1 year older customer has a 0.99 times shorter survival duration. \
    =\>Young customers are better!

-   The differences between customers coming from Google Ads versus Facebook ads are neglectable. There is a small tendency for Google Customers to have a 0.95 times lower customer-relationship, but this is not statistically significant. \
    =\> It doesn't matter where you advertise!

**🤴🏼Your best performing customer profile:**

-   High income

-   20-30 years old

**📈You should advertise on:**

-   Google or Facebook - it does not matter for churn
